Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/compilation.tex
more edits.

\para{Minimization}
\ratul{This minimization section can probably go entirely, shrunken, or moved to the implementation}
Before checking if policies captured by the PGIR are safe, we minimize the PGIR to better scale our failure safety checks. The minimization is based on the observation that although every path in the PGIR is a valid path in the topology, we do not want to consider loops when configuring the network. In particular, BGP's loop prevention mechanism, by which an AS rejects any route for which it is already in the AS path, means that loops should never occur.\footnote{ Misconfigured aggregation can still lead to loops in BGP.}
%
However, the PGIR may contain many paths that form loops in the topology. For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path that BGP announcements might take, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
Since we are primarily concerned with guaranteeing user policy correctness and safety with respect to network failures, it is useful to eliminate cases that are not possible due to loops. 
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.
We use graph dominators~\ref{bib:todo} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any simple path between the start and end nodes.
A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators in practice~\ref{bib:todo}.
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. 
For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
Colored nodes and dashed edges are removed after minimization since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

@@

To implement path preferences in routing, BGP uses local preferences on a per-device basis. However, the distributed nature of BGP makes setting preferences locally to achieve a network-route routing policy difficult, particularly in the presence of failures. For example, imagine an extremely simple policy for the topology in Figure~\ref{fig:example-compilation}, which says to prefer one route over another:
%
$$(W \cdot A \cdot C \cdot D) \gg (W \cdot B \cdot C \cdot E)$$
%
How could such a policy be implemented in BGP? Suppose we simply set the local preference at router $C$ to prefer $D$ over $E$. However, now if the A -- C link fails, then suddenly $C$ has made the wrong decision and can not get either the primary or backup route specified, despite the fact that the $W \cdot B \cdot C \cdot E$ path is available.

At a first cut, any time a router must make a decision locally between several route options, there is the possibility that it might choose incorrectly. PGIR captures this notion of choice precisely. For example, router $D$ in Figure~\ref{fig:example-compilation} appears only once in the PGIR, possibly receiving announcements from $X$ or $Y$. However, regardless of whether $D$ chooses a route from $X$ or $Y$, the set of paths allowed by the policy after going through $D$ will be the same in either case. This remains true despite any failures that might have occurred in the network. Thus $D$ can safely prefer $X$ and $Y$ equally.

@@

In general, ensuring that an individual router's preferences respect the policy under all possible failures is hard, and enumerating all failures is intractable. We thus adopt a conservative analysis based on the observations above. The high-level idea is to order each PGIR node according to a \textit{can prefer} relation. For example, node $(3,2) C$ can be preferred to node $(-,2) C$, but not the other way around. Intuitively, a node $N_1$ can be preferred to another node $N_2$ when, for each preference $N_2$ might achieve, there is a better preference that $N_1$ will achieve regardless of failures. Formally this means that $N_1$ can be preferred to $N_2$ when:
%
$$\forall i, \exists j, j \leq i \wedge protect(G_j, N_1, G_i, N_2)$$
%
where \textit{protect} means that, from $N_1$ on the PGIR restricted nodes that can potentially achieve preference $j$ or better, there are at least as many paths as from $N_2$ restricted to $G_i$

\begin{algorithm}[t!]
\caption{Failure Protection}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Protect($G_1$, $N_1$, $G_2$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \For {$x$ in $adj(G_2, n_2)$}
      \If {some $y$ in $adj(G_1,n_1)$ shadows $x$}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:failures} defines what it means for one node to \textit{protect} against the failures of another. It walks from nodes $N_1$ and $N_2$ in $G_j$ and $G_i$ respectively, and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, take an equivalent step. Thus any paths leading from $N_2$ will have an equivalent path over the same topology locations from $N_1$. Algorithm~\ref{alg:failures} works because each PGIR node can have at most one outgoing neighbor with the same topological location, so the next hop neighbors can be checked without backtracking. The algorithm terminates since the number of related states $(x,y)$ that can be expolored is finite.
\ryan{Add the new part that goes up to a dominating predecessor when no other option}

Local preferences are now obtained by, for each router in the topology, sorting the corresponding PGIR nodes according to \textit{can prefer} relation. If two nodes are incomparable, then the compiler rejects the policy as unimplementable.
\ryan{We are checking if the preorder forms a partial order by satisfying asymmetry}

@@

The checks for failure safety described above overlook a critical point: even if the analysis determines that a node $x$ protects against failures of another node $y$, it may be tricked into thinking $x$ has a particular path that will actually be discarded by BGP's loop prevention mechanism due to locations previously traversed before arriving at $x$. To avoid this situation, the compiler conservatively checks that $x$ protects against $y$ without using any nodes that shadow another node that appears above $x$ in the PGIR (i.e., nodes reachable from $x$ in $G_j^R$). It is fine however, to reuse the same nodes that appear above $x$, since any path that went through another node $z$ to get to $x$, but then looped through $z$ again, would be obtainable from $z$ itself.
\ryan{rework this}

@@

Once we have the total ordering on node preferences in the PGIR from the failure safety analysis, the translation to ABGP is straightforward. The idea is to encode the state of the automaton using BGP community values. Each router will match based on its peer and a community value corresponding to the state of the PGIR, and then update the state before exporting to the neighbors permitted by the PGIR. For example, router $A$ in Figure~\ref{fig:example-compilation} will allow an announcement from $C$ with a community value for state $(3,2)$ (and deny anything else). If it sees such an announcement, it will remove the old community value, and add a new one for state $(4,2)$ before exporting it to $W$.

@@

To make configurations more compact and readable, the \sysname compiler makes several passes over the ABGP policy, removing community tagging and matching when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. After removing the communities, the two rules at router $B$ are merged into a single rule.

>>>>

\para{Minimization}

Before checking if policies captured by the PGIR are safe, we minimize the PGIR to improve our safety analysis. 
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route for which it is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
It often simplifies the failure-safety analysis to eliminate cases that are not possible due to loops. 
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.
We use graph dominators~\ref{bib:todo} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading to $y$ from the start node. Many efficient algorithms exist for finding graph dominators in practice~\ref{bib:todo}.
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. 
For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
Colored nodes and dashed edges are removed after minimization since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

@@

To implement path preferences in routing, BGP uses local preferences on a per-device basis. However, the distributed nature of BGP makes setting preferences locally to achieve a network-route routing policy difficult, particularly in the presence of failures. For example, imagine an extremely simple policy for the topology in Figure~\ref{fig:example-compilation}, which says to prefer one route over another:
%
$$(W \cdot A \cdot C \cdot D) \gg (W \cdot B \cdot C \cdot E)$$
%
How could such a policy be implemented in BGP? Suppose we simply set the local preference at router $C$ to prefer $D$ over $E$. However, now if the A -- C link fails, then suddenly $C$ has made the wrong decision and can not get either the primary or backup route specified, despite the fact that the $W \cdot B \cdot C \cdot E$ path is available in the network.

At a first cut, any time a router must make a decision locally between several route options, there is the possibility that it might choose incorrectly. The PGIR captures this notion of choice precisely. For example, router $D$ in Figure~\ref{fig:example-compilation} appears only once in the PGIR, possibly receiving announcements from $X$ or $Y$. However, regardless of whether $D$ chooses a route from $X$ or $Y$, the set of paths allowed by the policy after going through $D$ will be the same in either case. This remains true despite any failures that might have occurred in the network. Thus $D$ can safely prefer $X$ and $Y$ equally.

@@

In general, ensuring that an individual router's preferences respect the policy under all possible failures is hard, and enumerating all failures is intractable. We thus adopt a conservative analysis based on the observations above. The high-level idea is to order each PGIR node according to a \textit{can prefer} relation. For example, node $(3,2) C$ can be preferred to node $(-,2) C$, but not the other way around. Intuitively, a node $N_1$ can be preferred to another node $N_2$ when, for each preference $N_2$ might achieve, there is a better preference that $N_1$ will achieve regardless of failures. Formally this means that $N_1$ can be preferred to $N_2$ when:
%
$$\forall i, \exists j, j \leq i \wedge protect(G_j, N_1, G_i, N_2)$$
%
where \textit{protect} means that, from $N_1$ on the product graph restricted to nodes that can potentially achieve preference $j$ or better, 
there is always a path to each each ending location whenever this is a path from $N_2$ restricted to $G_i$.

\begin{algorithm}[t!]
\caption{Failure Protection}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Protect($G_1$, $N_1$, $G_2$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \For {$x$ in adj($G_2$, $n_2$)}
      \If {\big($\exists y \in$ adj($G_1$,$n_1$), shadows $x$\big) \textbf{or} \\  
            \hspace{5.2em} \big($\exists y \in$ $G_1$, dominates $n_1$, shadows $x$\big)}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:failures} defines what it means for one node to \textit{protect} against the failures of another. It walks from nodes $N_1$ and $N_2$ in $G_j$ and $G_i$ respectively, and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, take an equivalent step. 
When there is no equivalent step, the algorithm checks to see if there is an equivalent dominator (a topology location the advertisement must have passed through earlier) that can protect against failures instead.
Thus any time $N_2$ can get to a final location for preference $i$, $N_1$ can get to at least the same final locations for a better preference $j$. The algorithm terminates since the number of related states $(x,y)$ that can be expolored is finite.

Local preferences are now obtained by, for each router in the topology, sorting the corresponding PGIR nodes according to \textit{can prefer} relation. If two nodes are incomparable, then the compiler rejects the policy as unimplementable.

@@

The checks for failure safety described above overlook one critical point: Paths for a higher preference might not actually exist due to loops. To avoid this situation, the compiler conservatively checks that, when node $x$ protects against failures for node $y$, its does so without using any nodes that shadow another node that appears ``above'' $x$ in the PGIR. 
%It is fine however, to reuse the same nodes that appear above $x$, since any path that went through another node $z$ to get to $x$, but then looped through $z$ again, would be obtainable from $z$ itself.
%even if the analysis determines that a node $x$ protects against failures of another node $y$, it may be tricked into thinking $x$ has a particular path that will actually be discarded by BGP's loop prevention mechanism due to locations previously traversed before arriving at $x$. To avoid this situation, the compiler conservatively checks that $x$ protects against $y$ without using any nodes that shadow another node that appears above $x$ in the PGIR (i.e., nodes reachable from $x$ in $G_j^R$). It is fine however, to reuse the same nodes that appear above $x$, since any path that went through another node $z$ to get to $x$, but then looped through $z$ again, would be obtainable from $z$ itself.

@@

Once we have the ordering on node preferences in the PGIR from the failure safety analysis, the translation to ABGP is straightforward. The idea is to encode the state of the automaton using BGP community values. Each router will match based on its peer and a community value corresponding to the state of the PGIR, and then update the state before exporting to the neighbors permitted by the PGIR. For example, router $A$ in Figure~\ref{fig:example-compilation} will allow an announcement from $C$ with a community value for state $(3,2)$ (and deny anything else). If it sees such an announcement, it will remove the old community value, and add a new one for state $(4,2)$ before exporting it to $W$.

@@

To make configurations more compact and readable, the \sysname compiler makes several passes over the ABGP policy, removing community tagging and matching when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.
